\documentclass{book}

\input{../activities-preamble.tex}
\begin{document}
\setcounter{project}{173}
\addtocounter{project}{-1}
\begin{activity}[]\label{act-hdseq}
\hypertarget{p-1008}{}%
\(2n\) people stand in line at a old-timey movie theater. Admission is 50 cents, (denoted by H), and the box office starts with no change. \(n\) of the people have H and \(n\) have \textdollar{}1 (denoted D). In how many ways can the \(2n\) people line up so that all can be admitted?%
\par\smallskip%
\noindent\textbf{Hint.}\hypertarget{hint-113}{}\quad%
\hypertarget{p-1009}{}%
Here we enumerate the number of workable sequences of \(n\) H's and \(n\) D's such that at each point in the sequence the number of H's is not less than the number of D's.  Try writing out all HD sequences and see which ones are acceptable and which are not.%
~\hfill{\tiny\hyperlink{a-173}{[hint]}\hypertarget{q-173}{}}\par\smallskip%
\noindent\textbf{Solution.}\hypertarget{solution-83}{}\quad%
\hypertarget{p-1010}{}%
The total number of sequences of \(n\) H's and \(n\) D's is \(\binom{2n}{n}\). We delete the number of \emph{nonworkable} sequences. Each such sequence has a first snag as shown by the highlighted D in%
\begin{equation*}
\text{H\ H\ D\ D\ \alert{D}\ H\ H\ D\ D\ D\ H\ H}
\end{equation*}
Reverse each letter up to and including the snag, obtaining%
\begin{equation*}
\text{D D H H H H H D D D H H},
\end{equation*}
a sequence of \(n+1\) H's and \(n-1\) D's. For \(n = 3\) , H D \alert{D} D H H is a non-workable sequence with the snag indicated. Its mate, found by reversing the first three letters, is D H H D H H, a sequence of 2D's and 4H's. \emph{Any} arrangement of 2D's and 4H's will correspond to exactly one non-workable sequence; simply scan through and see the \emph{first} time the H's dominate the D's and then reverse through that spot. There are \(\binom{6}{3}
= 20\) arrangements of 3H's and 3D's. There are \(\binom{6}{2}
= 15\) arrangements of 2D's and 4H's each of which corresponds to a non-workable sequence. Hence there are \(\binom{6}{3}
-\binom{6}{2} = 5\) workable sequences. In general, \(\binom{2n}{n}  - \binom{2n}{n - 1} = \frac{1}{n + 1}\binom{2n}{n}\) gives the number of workable sequences.%
\end{activity}

\clearpage\end{document}
