\documentclass{book}

\input{../activities-preamble.tex}
\begin{document}
\setcounter{project}{46}
\addtocounter{project}{-1}
\begin{activity}[]\label{activity-39}
\hypertarget{p-411}{}%
Prove by induction on vertices that any graph \(G\) which contains at least one vertex of degree less than \(\Delta(G)\) (the maximal degree of all vertices in \(G\)) has chromatic number at most \(\Delta(G)\).%
\par\smallskip%
\noindent\textbf{Solution.}\hypertarget{solution-35}{}\quad%
\begin{proof}\hypertarget{proof-5}{}
\hypertarget{p-412}{}%
Let \(G\) be a graph with \(n\) vertices, maximal degree \(\Delta(G)\) and at least one vertex of degree less than \(\Delta(G)\). Assume for the sake of induction that all graphs \(G'\) with fewer than \(n\) vertices and a vertex of degree less than \(\Delta(G')\) have chromatic number less than \(\Delta(G')\).%
\par
\hypertarget{p-413}{}%
Find a vertex of \(G\) with degree less than \(\Delta(G)\) and remove it. This forms a subgraph \(G'\) which has \(n-1\) vertices. Also, since we removed edges, we know that \(\Delta(G') \le \Delta(G)\). If these maximal degrees are equal, then \(G'\) must also have a vertex of degree less than \(\Delta(G')\), since at least one of its vertices had one more edge in \(G\). In this case, we can apply our inductive hypothesis to produce a coloring of the vertices of \(G'\) using just \(\Delta(G)\) colors. If \(\Delta(G') \lt  \Delta(G)\), then we also can easily find a proper coloring of the vertices of \(G'\) using just \(\Delta(G)\) colors by starting with any vertex and coloring it and all of its neighbors differently, and then fanning out. Thus \(G'\) has a proper vertex coloring using just \(\Delta(G)\)-many colors.%
\par
\hypertarget{p-414}{}%
Now move back to \(G\). Put the removed vertex back into the graph. Since it is adjacent to at most \(\Delta(G) - 1\) other vertices, there will be one of the \(\Delta(G)\) colors that is not present among its neighbors, which we could use to color the newly inserted vertex.%
\end{proof}
\end{activity}

\clearpage\end{document}
