\documentclass{book}

\input{../activities-preamble.tex}
\begin{document}
\setcounter{project}{43}
\addtocounter{project}{-1}
\begin{activity}[]\label{activity-36}
\leavevmode%
\begin{enumerate}[font=\bfseries,label=(\alph*),ref=\alph*]
\item\label{task-58} \hypertarget{p-402}{}%
Find a graph that has chromatic number 3 even though it does not contain \(K_3\) as a subgraph.%
\par\smallskip%
\noindent\textbf{Solution.}\hypertarget{solution-42}{}\quad%
\hypertarget{p-403}{}%
The graph \(C_5\) (a cycle of length 5) has chromatic number 3, but does not have a triangle.%
\item\label{task-59} \hypertarget{p-404}{}%
Find a graph that has chromatic number 4 even though it does not contain \(K_4\) as a subgraph.%
\par\smallskip%
\noindent\textbf{Hint.}\hypertarget{hint-16}{}\quad%
\hypertarget{p-405}{}%
Modify your \(K_3\)-free graph with chromatic number 3.%
~\hfill{\tiny\hyperlink{a-43.b}{[hint]}\hypertarget{q-43.b}{}}\par\smallskip%
\noindent\textbf{Solution.}\hypertarget{solution-43}{}\quad%
\hypertarget{p-406}{}%
The graph below has chromatic number 4, but does not contain a copy of \(K_4\).  The 5-cycle requires 3 colors, and the vertex in the center is adjacent to all of these, so requires a 4th color.  These graphs are sometimes called \terminology{wheel graphs}.%
\begin{sidebyside}{1}{0.4}{0.4}{0}
\begin{sbspanel}{0.2}
\resizebox{\linewidth}{!}{{
 \begin{tikzpicture}
   \foreach \x in {0,...,4}{
 \draw (0,0) -- (90+\x*360/5:1) \v -- (162+\x*360/5:1);
}
\draw (0,0) \v;
 \end{tikzpicture}
}
}
\end{sbspanel}
\end{sidebyside}
\item\label{task-60} \hypertarget{p-407}{}%
Challenge: Find a graph that has chromatic number 4 even though it does not contain \(K_3\) as a subgraph.%
\par\smallskip%
\noindent\textbf{Hint.}\hypertarget{hint-17}{}\quad%
\hypertarget{p-408}{}%
The smallest such graph contains 11 vertices.  Try starting with \(C_5\) and adding vertices creating a lot of odd cycles.%
~\hfill{\tiny\hyperlink{a-43.c}{[hint]}\hypertarget{q-43.c}{}}\par\smallskip%
\noindent\textbf{Solution.}\hypertarget{solution-44}{}\quad%
\hypertarget{p-409}{}%
The graph below has chromatic number 4 but does not contain \(K_3\).  This is called the Grötzsch graph.%
\begin{sidebyside}{1}{0.35}{0.35}{0}
\begin{sbspanel}{0.3}
\resizebox{\linewidth}{!}{{
 \begin{tikzpicture}
   \foreach \x in {0,...,4}{
\draw (0,0) \v -- (90+\x*72:1) \v -- (162+\x*72:2) \v -- (90+\x*72:2) -- (162+\x*72:1);
}
 \end{tikzpicture}
}
}
\end{sbspanel}
\end{sidebyside}
\end{enumerate}
\end{activity}
\end{document}
